State of the Practice for Lattice Boltzmann Method Software

### 2022-01-07 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Further edits related to common practices

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 103 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP), a multi-criteria
    decision making method appropriate for cases with a mix of qualitative and
    quantitative factors. Our ranking was roughly consistent with ranking the
    repositories by the number of GitHub stars, although the number one project
    by star count (Sailfish) is ranked 9th by our methodology. Our top three
    packages were ESPResSo, Ludwig, and Palabos (all in the top 10 by star
    count). We interviewed 4 developers to gain insight into their current pain
    points. Identified challenges include lack of development time, lack of
    funding, and difficulty with ensuring correctness.  Based on our assessment,
    we make the following recommendations for current and future LBM projects:
    provide a detailed user manual and tutorials, explicitly state the limits of
    the software, use user-friendly programming languages, consider a peer
    review process, communicate development and contribution information, and
    use continuous integration and project management tools.

@@

The candidate software was found through search engine queries targeting
authoritative lists of software. We found LBM software listed on the websites
GitHub and swMATH, as well as through articles found in scholarly journals and
databases.  The Domain Expert (Section~\ref{domainexpertrecommend}) was also
engaged in selecting the candidate software.  \wss{Packages that fit our
criteria but were found after data collection and analysis was conducted, such
as Musubi \citep{HasertEtAl2014}, will need to be added to future SOP
assessments.}

@@

Another raised concern was that there are no organizations helping with the
development of good quality software; but some do exist, including
\href{https://bssw.io/}{Better Scientific Software (BSSw)},
\href{https://www.software.ac.uk/}{Software Sustainability Institute}, and
\href{https://software-carpentry.org/}{Software Carpentry}. Some SCS developers
may not be familiar with these organizations.

@@

\subsection{Current Practices} \label{currentPractices}

The following points regarding software quality should be considered when
developing LBM software packages. These points are based on developer
interviews, SCS literature, what was found to have worked for packages that were
designated as high quality in this assessment.  Our recommendations are not
lists of what should have been done in the past, or what should be done now;
they are just suggestions for consideration in the future.  \wss{Rework this
section to highlight what the top projects are doing.  Point to checklists for
other advice?}

\begin{itemize}
	\item Surface Reliability. Include descriptive error messages where
	appropriate. This was done by most packages that encountered a fault.
	\item Surface Reliability. In case automatic installation of dependencies
	fails, the system should indicate to the user what dependencies need to be
	installed manually. This was done by many packages, including ESPResSo++,
	Ludwig, LUMA, pyLBM, TCLB.
	\item Surface Reliability. The packages should include detailed tutorials,
	including dependencies, expected output, and any additional supplementary
	documentation that may be required. This was done by many packages,
	including waLBerla, Palabos, MechSys, LUMA, and pyLBM.  \wss{Comment from
	ASCR attendee that success with open source requires good tutorials and
	documentation - is this published anywhere?}
	\item Surface Robustness. The system must provide descriptive error messages
	when it encounters unexpected input, including incorrect data types, empty
	input, and missing files or links. Nineteen of the software packages behaved
	reasonably when tested with unexpected input.
	\item Surface Performance. Integrate parallelization tools and techniques to
	reduce processing time. Consider GPU processing, CUDA, and MPI. Eighteen
	software packages mentioned parallelization in their artifacts.
	\item Surface Performance. The user should be able to choose to process
	their model on either the CPU or GPU. ESPResSo, lbmpy, lettuce, pyLBM, and
	TCLB allow for either CPU or GPU processing.
	\item Surface Usability. Include user hardware and software requirements
	documentation. Hardware requirements are rarely listed in the software
	packages that were assessed. Only those offering GPU processing (ESPResSo,
	lbmpy, lettuce, pyLBM, Sailfish, TCLB) mentioned any hardware requirements.
	On the other hand, some sort of software requirements were available for all
	packages, even if only some dependencies or a compatible operating system
	were mentioned.
	\item Surface Usability. Provide a user tutorial that indicates the expected
	output. This was done by many packages, including waLBerla, Palabos,
	MechSys, LUMA, pyLBM.
	\item Surface Usability. Provide a detailed user manual. It should identify
	elements of user interfaces, and identify all requirements to model a
	system. Thirteen packages have a user manual. ESPResSo has a well detailed
	manual.
	\item Maintainability. Ideally include most of the common and less common
	artifacts listed in Section~\ref{artifacts}.
	\item Surface Understandability. Use consistent, distinctive, and meaningful
	code identifiers. Twenty of the 24 software packages use such code
	identifiers.
	\item Surface Understandability. Use symbolic constants. Thirteen of the 24
	software packages use such constants.
	\item Surface Understandability. Identify algorithms that are used in source
	code comments. Document them in the artifacts. Cite relevant external
	sources if needed. Domain algorithms are noted in the source code of 11
	packages.
	\item Surface Understandability. Provide a user manual that identifies all
	technical and problem modeling requirements. Thirteen software packages have
	a user manual.
	\item Unambiguity. Provide a detailed user manual. Thirteen packages have a
	user manual. ESPResSo has a well detailed manual.

@@

>>>>

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 103 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP), a multi-criteria
    decision making method appropriate for cases with a mix of qualitative and
    quantitative factors. Our ranking was roughly consistent with ranking the
    repositories by the number of GitHub stars, although the number one project
    by star count (Sailfish) is ranked 9th by our methodology. Our top three
    packages were ESPResSo, Ludwig, and Palabos (all in the top 10 by star
    count). We interviewed 4 developers to gain insight into their current pain
    points. Identified challenges include lack of development time, lack of
    funding, and difficulty with ensuring correctness.  Based on our assessment,
    we make the following recommendations for current and future LBM projects:
    provide a detailed user manual and tutorials, explicitly state the limits of
    the software, use user-friendly programming languages, consider a peer
    review process, communicate development and contribution information, and
    use continuous integration and project management tools.  \wss{Change the
    abstract to discuss how LBM compares to common practices and best practices.
    Highlight some future practices.}

@@

The candidate software was found through search engine queries targeting
authoritative lists of software. We found LBM software listed on the websites
GitHub and swMATH, as well as through articles found in scholarly journals and
databases.  The Domain Expert (Section~\ref{sec_vet_software_list}) was also
engaged in selecting the candidate software.  \wss{Packages that fit our
criteria but were found after data collection and analysis was conducted, such
as Musubi \citep{HasertEtAl2014}, will need to be added to future SOP
assessments.}

@@

Another raised concern was that there are no organizations helping with the
development of good quality software; but some do exist, including
\href{https://bssw.io/}{Better Scientific Software (BSSw)},
\href{https://www.software.ac.uk/}{Software Sustainability Institute} \citep{CrouchEtAl2013}, and
\href{https://software-carpentry.org/}{Software Carpentry}. Some SCS developers
may not be familiar with these organizations. \citep{WilsonAndLumsdaine2006}

@@

\subsection{Common Practices} \label{currentPractices}

The practices that arose frequently.  Compare and contrast to what is done with
research software in general.  The following points regarding software quality
should be considered when developing LBM software packages. These points are
based on developer interviews, SCS literature, what was found to have worked for
packages that were designated as high quality in this assessment.  Our
recommendations are not lists of what should have been done in the past, or what
should be done now; they are just suggestions for consideration in the future.
\wss{Rework this section to highlight what the top projects are doing.  Point to
checklists for other advice?}  Highlights of common practices.  Put in context.

Conclusion that LBM software common practice is similar to the common practice
elsewhere.  Maybe point out where LBM is ahead.  Use results from common
artifacts, common process etc. to support conclusion.  (Continuous integration, version control, README, Req etc. - point to Table 8.)

\begin{itemize}
	\item Surface Reliability. Surface Usability. The packages should include
	detailed tutorials, including dependencies, expected output, and any
	additional supplementary documentation that may be required. This was done
	by many packages, including waLBerla, Palabos, MechSys, LUMA, and pyLBM.
	\wss{Comment from ASCR attendee that success with open source requires good
	tutorials and documentation - is this published anywhere?  I think I can
	find this recommendation in other papers.}
	\item Surface Usability. Unambiguity. Surface Understandability. Provide a detailed user manual. It
	should identify elements of user interfaces, and identify all requirements
	to model a system. Thirteen packages have a user manual. ESPResSo has a well
	detailed manual. \wss{Include this with tutorial?  Find papers that
	recommend this kind of documentation?}
	\item Surface Usability. Include user hardware and software requirements
	documentation. Hardware requirements are rarely listed in the software
	packages that were assessed. Only those offering GPU processing (ESPResSo,
	lbmpy, lettuce, pyLBM, Sailfish, TCLB) mentioned any hardware requirements.
	On the other hand, some sort of software requirements were available for all
	packages, even if only some dependencies or a compatible operating system
	were mentioned.  \wss{This is a key part of reproducibility.}
    \item Not much of a process - find papers that support this - papers recommend "Define clear and transparent contribution, governance and communication processes" JimenezEtAl2017
\end{itemize}

@@

- introduce the idea of using our measurement template as a checklist, instead of, or in addition to other checklists - cite the other checklists \citep{SchlauchEtAl2018, BrettEtAl2021, vanGompelEtAl2016}, \href{Software Sustainability Institute Online Form}{https://docs.google.com/forms/d/e/1FAIpQLSf0ccsVdN-nXJCHLluJ-hANZlp8rDKgprJa0oTYiLZSDxh3DA/viewform}

- discuss Wilson papers ``Best Practices'' \citep{WilsonEtAl2014} and ``Good Enough Practices'' \citep{WilsonEtAl2016}

- mention flake8 for Python?

- mention \href{https://joss.theoj.org/}{The Journal of Open Source Software}

- reproducibility - can point to the PiccoloAndFrampton2016 paper

- AlNoamanyAndBorghi2018 lists the most popular languages, Python is number 1

- JimenezEtAl2017 talks about Define clear and transparent contribution, governance and communication processes
